Parsing Tweets into Universal Dependencies

### 2018-04-16 paper.tex
update

%\newcommand{\yjcomment}[1]{\textcolor{orange}{[$_\mathrm{L}^\mathrm{Y}$#1]}}
%\newcommand{\nascomment}[1]{\textcolor{blue}{[#1 ---\textsc{nas}]}}
%\newcommand{\yicomment}[1]{\textcolor{gray}{[#1 ---\textsc{yi}]}}
\newcommand{\nss}[1]{\textcolor{magenta}{[$_\mathrm{S}^\mathrm{NS}$#1]}}

@@

\begin{itemize}
%\item For the at-mentions in retweet constructions, we attach them to their retweet markers. We will discuss our choice in the following paragraph;
\item If a non-syntactic token is within a sentence that has a clear predicate, it will be attached to this predicate. 
The retweet construction is a special case and we will discuss its treatment in the following paragraph.
\item If the whole sentence\nss{tweet?} is a sequence of non-syntactic tokens, we attach all these tokens to the first one.
%\item In parataxis case, when it is difficult to decide which sentence or phrase x should belong to, we always attach x to the previous predicate of the paratactic clause or phrase;
\item Non-syntactic tokens are mostly labeled as {\it discourse}, but
  URLs are always labeled as {\it list}, following the UD\_English-EWT dataset.% \yicomment{need to add punct for punctuations?}
%\\ \yicomment{For retweet at-mention case, we attach the retweet at-mentions to the RT sign, so that the retweet action and the people whose tweet is being retweeted could be connected.\footnote{We can also follow the first convention to attach the retweet at-mention to its predicate, but we think it is better for downstream tasks.}}
\end{itemize}

@@

%In the syntax annotation, tweet specific
%constructions are considered by both \citet{sanguinetti-17} and our paper.
With respect to tweet-specific constructions,
\citeposs{sanguinetti-17} and our  interpretation of headedness is the same, 
but we differ in the relation label.
For topical hashtags, we use \textit{discourse} while \citet{sanguinetti-17} used \textit{parataxis}. 
In referential URLs, we use \textit{list} (following the precedent of UD\_English-EWT) while \citet{sanguinetti-17} used \textit{dep}.
%<<<<<<< HEAD
%The choice of relation on the non-syntactic tokens is questionable
%since UD doesn't specify a convention for this. 
%Our choice of \textit{discourse} is inspired by the observation that emoticon is annotated as \textit{discourse}.
%\citet{sanguinetti-17} used the same relation for the emoticon but analyze topical hashtag as \textit{parataxis}.%\nss{what did PosTWITA-UD use for non-syntactic tokens?}\yjcomment{done}
%=======
%The choice of relation on the non-syntactic tokens is questionable
%since UD doesn't specify a convention for this. 
Our choice of \textit{discourse} for sentiment emoticons
is inspired by the observation that emoticons are annotated as \textit{discourse} by UD\_English-EWT
and \citet{sanguinetti-17} used the same relation for the emoticons.\nss{how does the other treebank treat the RT construction and truncated words?}
%\nss{topical hashtags already mentioned above}
%>>>>>>> a7f86b777a73f64b2f05101f4804421aea0412f3
% Non-syntactic tokens can be easily recovered
% from our annotation and changing it to other relation in the future development is not difficult.
We expect that the official UD guidelines will eventually adopt standards for these constructions so the treebanks can be harmonized.

@@

\subsubsection{Data Collection}
{\sc Tweebank v2} is built on the original data of {\sc Tweebank v1}
(840 unique tweets, 639/201 for training/test), along with an 
additional 210 tweets sampled from the POS-tagged dataset of
\newcite{gimpel-EtAl:2011:ACL-HLT2011} and 2,500 tweets sampled
from the Twitter stream from February 2016 to July 2016.\footnote{Data downloaded from \url{https://archive.org/}.}
The latter data source consists of 147.4M English tweets after being
filtered by the {\it lang} attribute in the tweet JSON and {\it
  langid.py}.\footnote{\url{https://github.com/saffsd/langid.py}}
In the same way as \citet{kong-EtAl:2014:EMNLP2014},
reference unit is always the tweet in its entirety
-- which may thus consist of multiple sentences -- not the sentence alone.\nss{``Each tweet is annotated with a single tree, rather than being split into sentences.''?}
Before annotation, we use simple regular expression to anonymize usernames and URLs.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
%		\hline
		\it System & $F_1$ \\
		\hline
		  Stanford CoreNLP & 97.3 \\
		 Twokenizer & 94.6 \\
% 		UDPipe v1.2 & 94.7 \\
		\hdashline
		UDPipe v1.2 & 97.4 \\
	 	our bi-LSTM tokenizer & 98.3 \\
%		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}
%\citet{wang-EtAl:2017:Long6} show that the parser\nss{tokenizer?}\yjcomment{it's a parsing work.}\nss{I guess I don't follow how parsing results 
%on combined training data affect our decision about how to tokenize.} trained 
%on the combination of the UD\_English and Singapore English dialect (Singlish) treebanks
%outperforms the one trained only on the dialect treebank.  
Our preliminary results
showed that our model trained on
the combination of UD\_English-EWT and \textsc{Tweebank v2}
outperformed the one trained only on the UD\_English-EWT or \textsc{Tweebank v2},
consistent with previous work on dialect treebank parsing \cite{wang-EtAl:2017:Long6}.
So we trained our tokenizer on the training portion of
\textsc{Tweebank v2} combined with the UD\_English-EWT training set
and tested on the \textsc{Tweebank v2} test set. % \yicomment{should mention that combination of corpora improves the results, otherwise we should answer why we didn't try training ONLY on Tweebank v2}\yjcomment{add discussion}
We report $F_1$ scores, combining precision and recall for token identification. Table \ref{tbl:tok-result} shows the
tokenization results, compared to  other available tokenizers. 
 Stanford CoreNLP \cite{manning-EtAl:2014:P14-5} and Twokenizer
\cite{ICWSM101540} are rule-based systems and were not adapted
to the UD tokenization scheme.
The UDPipe v1.2
\cite{straka-strakova:2017:K17-3} model was re-trained on the same
data as our system. Compared with UDPipe, we use an LSTM
instead of a GRU in our model and we also use a larger size for hidden units (64 vs.~20),
which has stronger representational power. 
Our bi-LSTM tokenizer achieves the best accuracy among all these
tokenizers.  These results speak to the value of statistical modeling
in tokenization for informal texts.%\nss{this paragraph needs copyediting}

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
%		\hline
		\it System & Precision \\
		\hline
%		Stanford Postagger & 65.6 \\
		Stanford CoreNLP & 90.6 \\
%		UDPipe v1.2 & 78.7 \\
		\citealp{owoputi-EtAl:2013:NAACL-HLT} (greedy) & 93.7 \\
		\citealp{owoputi-EtAl:2013:NAACL-HLT} (CRF) & 94.6 \\
		\hdashline
%		UDPipe v1.2 & 83.8 \\
		\citealp{ma-hovy:2016:P16-1} & 92.5 \\
%		our word bi-LSTM & 89.2 \\
%		our character + word bi-LSTMs & 91.5 \\
%		\hline
	\end{tabular}
	\caption{POS tagger comparison on gold-standard tokens in the
          \textsc{Tweebank v2} test set.\nss{why Precision and not Accuracy?} \label{tbl:pos-result}}
\end{table}

@@

In our preliminary\nss{initial?} experiments, we train our parser on the combination of UD\_English-EWT
and \textsc{Tweebank v2} training sets. Gold-standard tokenization and automatic POS
tags are used. Automatic POS tags are assigned with 5-fold
jackknifing. Hyperparameters %, including the value of $\alpha$ in Equation \ref{eq:distill},
are tuned on the \textsc{Tweebank v2} development set. Unlabeled attachment score and
labeled attachment score (including punctuation) are reported.
All the experiments were run on a Xeon E5-2670 2.6 GHz machine.

@@

\paragraph{Learning from exploration.} When we set $\alpha =1$, we
eliminate the oracle from the estimation procedure (for the distilled
model).  This presents an opportunity to learn with \emph{exploration}, by
randomly sampling transitions from the ensemble, found useful
in recent methods for training greedy models that use dynamic oracles
\citep{goldberg-nivre:2012:PAPERS,TACL145,TACL885,ballesteros-EtAl:2016:EMNLP2016}.  
We find that this
approach outperforms the  conventional distillation model, coming in
only one\nss{UAS? LAS is 1.5 points} point behind the ensemble (last line of Table~\ref{tbl:parse-result}).

@@

\paragraph{Pipeline evaluation.} Finally, we report  our full
pipeline's performance in  Table \ref{tbl:pipline}. We also compare
our model with a pipeline of the state-of-the-art systems (labeled  ``SOTA''):
Stanford CoreNLP tokenizer,\footnote{We choose the Stanford CoreNLP tokenizer in the spirit of comparing 
rule-based and statistical methods.} %\yicomment{why not compare with UDPipe?}\yjcomment{i was thinking stanford tokenizer is more suitable, but i will get the udpipe result},
\citeposs{owoputi-EtAl:2013:NAACL-HLT} tagger, and \citeposs{dozat-qi-manning:2017:K17-3} parser.
Our system differs from the SOTA pipeline in the tokenization and parser components.
From Table \ref{tbl:pipline}, our pipeline outperforms the SOTA when
evaluated in pipeline manner.
The results also emphasize the
importance of tokenization:  without gold tokenization
%\nascomment{changed ``segmentation'' to ``tokenization'' here, twice},
UD parsing
performance drops by more than\nss{? 77.9 vs. 74.0, right?} four points.

>>>>

%\newcommand{\yjcomment}[1]{\textcolor{orange}{[$_\mathrm{L}^\mathrm{Y}$#1]}}
%\newcommand{\nascomment}[1]{\textcolor{blue}{[#1 ---\textsc{nas}]}}
%\newcommand{\yicomment}[1]{\textcolor{gray}{[#1 ---\textsc{yi}]}}
%\newcommand{\nss}[1]{\textcolor{magenta}{[$_\mathrm{S}^\mathrm{NS}$#1]}}

@@

\begin{itemize}
%\item For the at-mentions in retweet constructions, we attach them to their retweet markers. We will discuss our choice in the following paragraph;
\item If a non-syntactic token is within a sentence that has a clear predicate, it will be attached to this predicate. 
The retweet construction is a special case and we will discuss its treatment in the following paragraph.
\item If the whole sentence is a sequence of non-syntactic tokens, we attach all these tokens to the first one.
%\item In parataxis case, when it is difficult to decide which sentence or phrase x should belong to, we always attach x to the previous predicate of the paratactic clause or phrase;
\item Non-syntactic tokens are mostly labeled as {\it discourse}, but
  URLs are always labeled as {\it list}, following the UD\_English-EWT dataset.% \yicomment{need to add punct for punctuations?}
%\\ \yicomment{For retweet at-mention case, we attach the retweet at-mentions to the RT sign, so that the retweet action and the people whose tweet is being retweeted could be connected.\footnote{We can also follow the first convention to attach the retweet at-mention to its predicate, but we think it is better for downstream tasks.}}
\end{itemize}

@@

%In the syntax annotation, tweet specific
%constructions are considered by both \citet{sanguinetti-17} and our paper.
With respect to tweet-specific constructions,
\citeposs{sanguinetti-17} and our  interpretation of headedness is the same, 
but we differ in the relation label.
For topical hashtags, we use \textit{discourse} while \citet{sanguinetti-17} used \textit{parataxis}. 
In referential URLs, we use \textit{list} (following the precedent of UD\_English-EWT) while \citet{sanguinetti-17} used \textit{dep}.
%<<<<<<< HEAD
%The choice of relation on the non-syntactic tokens is questionable
%since UD doesn't specify a convention for this. 
%Our choice of \textit{discourse} is inspired by the observation that emoticon is annotated as \textit{discourse}.
%\citet{sanguinetti-17} used the same relation for the emoticon but analyze topical hashtag as \textit{parataxis}.%\nss{what did PosTWITA-UD use for non-syntactic tokens?}\yjcomment{done}
%=======
%The choice of relation on the non-syntactic tokens is questionable
%since UD doesn't specify a convention for this. 
Our choice of \textit{discourse} for sentiment emoticons
is inspired by the observation that emoticons are annotated as \textit{discourse} by UD\_English-EWT
and \citet{sanguinetti-17} used the same relation for the emoticons.
Retweet constructions and truncated words were not explicitly touched
in \citet{sanguinetti-17}.
Judging from the released treebank\footnote{\url{https://github.com/UniversalDependencies/UD_Italian-PoSTWITA}}, the \textit{RT} marker, at-mention, and colon in the retweet construction are all attached to the predicate of the following sentence with \textit{dep}, \textit{vocative:mention} and \textit{punct}.
%\nss{how does the other treebank treat the RT construction and truncated words?}
%\nss{topical hashtags already mentioned above}
%>>>>>>> a7f86b777a73f64b2f05101f4804421aea0412f3
% Non-syntactic tokens can be easily recovered
% from our annotation and changing it to other relation in the future development is not difficult.
We expect that the official UD guidelines will eventually adopt standards for these constructions so the treebanks can be harmonized.

@@

\subsubsection{Data Collection}
{\sc Tweebank v2} is built on the original data of {\sc Tweebank v1}
(840 unique tweets, 639/201 for training/test), along with an 
additional 210 tweets sampled from the POS-tagged dataset of
\newcite{gimpel-EtAl:2011:ACL-HLT2011} and 2,500 tweets sampled
from the Twitter stream from February 2016 to July 2016.\footnote{Data downloaded from \url{https://archive.org/}.}
The latter data source consists of 147.4M English tweets after being
filtered by the {\it lang} attribute in the tweet JSON and {\it
  langid.py}.\footnote{\url{https://github.com/saffsd/langid.py}}
In the same way as \citet{kong-EtAl:2014:EMNLP2014},
reference unit is always the tweet in its entirety
-- which may thus consist of multiple sentences -- not the sentence alone.%\nss{``Each tweet is annotated with a single tree, rather than being split into sentences.''?}
Before annotation, we use a simple regular expression to anonymize usernames and URLs.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
%		\hline
		\it System & $F_1$ \\
		\hline
		  Stanford CoreNLP & 97.3 \\
		 Twokenizer & 94.6 \\
% 		UDPipe v1.2 & 94.7 \\
		\hdashline
		UDPipe v1.2 & 97.4 \\
	 	our bi-LSTM tokenizer & 98.3 \\
%		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}
%\citet{wang-EtAl:2017:Long6} show that the parser\nss{tokenizer?}\yjcomment{it's a parsing work.}\nss{I guess I don't follow how parsing results 
%on combined training data affect our decision about how to tokenize.} trained 
%on the combination of the UD\_English and Singapore English dialect (Singlish) treebanks
%outperforms the one trained only on the dialect treebank.  
Our preliminary results
showed that our model trained on
the combination of UD\_English-EWT and \textsc{Tweebank v2}
outperformed the one trained only on the UD\_English-EWT or \textsc{Tweebank v2},
consistent with previous work on dialect treebank parsing \cite{wang-EtAl:2017:Long6}.
So we trained our tokenizer on the training portion of
\textsc{Tweebank v2} combined with the UD\_English-EWT training set
and tested on the \textsc{Tweebank v2} test set. % \yicomment{should mention that combination of corpora improves the results, otherwise we should answer why we didn't try training ONLY on Tweebank v2}\yjcomment{add discussion}
We report $F_1$ scores, combining precision and recall for token identification. Table \ref{tbl:tok-result} shows the
tokenization results, compared to  other available tokenizers. 
 Stanford CoreNLP \cite{manning-EtAl:2014:P14-5} and Twokenizer
\cite{ICWSM101540}\footnote{We use the updated version of Twokenizer from \citet{owoputi-EtAl:2013:NAACL-HLT}.} are rule-based systems and were not adapted
to the UD tokenization scheme.
The UDPipe v1.2
\cite{straka-strakova:2017:K17-3} model was re-trained on the same
data as our system. Compared with UDPipe, we use an LSTM
instead of a GRU in our model and we also use a larger size for hidden units (64 vs.~20),
which has stronger representational power. 
Our bi-LSTM tokenizer achieves the best accuracy among all these
tokenizers.  These results speak to the value of statistical modeling
in tokenization for informal texts.%\nss{this paragraph needs copyediting}

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
%		\hline
		\it System & Accuracy \\
		\hline
%		Stanford Postagger & 65.6 \\
		Stanford CoreNLP & 90.6 \\
%		UDPipe v1.2 & 78.7 \\
		\citealp{owoputi-EtAl:2013:NAACL-HLT} (greedy) & 93.7 \\
		\citealp{owoputi-EtAl:2013:NAACL-HLT} (CRF) & 94.6 \\
		\hdashline
%		UDPipe v1.2 & 83.8 \\
		\citealp{ma-hovy:2016:P16-1} & 92.5 \\
%		our word bi-LSTM & 89.2 \\
%		our character + word bi-LSTMs & 91.5 \\
%		\hline
	\end{tabular}
	\caption{POS tagger comparison on gold-standard tokens in the
          \textsc{Tweebank v2} test set.%\nss{why Precision and not Accuracy?}
          \label{tbl:pos-result}}
\end{table}

@@

In our initial experiments, we train our parser on the combination of UD\_English-EWT
and \textsc{Tweebank v2} training sets. Gold-standard tokenization and automatic POS
tags are used. Automatic POS tags are assigned with 5-fold
jackknifing. Hyperparameters %, including the value of $\alpha$ in Equation \ref{eq:distill},
are tuned on the \textsc{Tweebank v2} development set. Unlabeled attachment score and
labeled attachment score (including punctuation) are reported.
All the experiments were run on a Xeon E5-2670 2.6 GHz machine.

@@

\paragraph{Learning from exploration.} When we set $\alpha =1$, we
eliminate the oracle from the estimation procedure (for the distilled
model).  This presents an opportunity to learn with \emph{exploration}, by
randomly sampling transitions from the ensemble, found useful
in recent methods for training greedy models that use dynamic oracles
\citep{goldberg-nivre:2012:PAPERS,TACL145,TACL885,ballesteros-EtAl:2016:EMNLP2016}.  
We find that this
approach outperforms the  conventional distillation model, coming in
1.5 point behind the ensemble (last line of Table~\ref{tbl:parse-result}).

@@

\paragraph{Pipeline evaluation.} Finally, we report  our full
pipeline's performance in  Table \ref{tbl:pipline}. We also compare
our model with a pipeline of the state-of-the-art systems (labeled  ``SOTA''):
Stanford CoreNLP tokenizer,\footnote{We choose the Stanford CoreNLP tokenizer in the spirit of comparing 
rule-based and statistical methods.} %\yicomment{why not compare with UDPipe?}\yjcomment{i was thinking stanford tokenizer is more suitable, but i will get the udpipe result},
\citeposs{owoputi-EtAl:2013:NAACL-HLT} tagger, and \citeposs{dozat-qi-manning:2017:K17-3} parser.
Our system differs from the SOTA pipeline in the tokenization and parser components.
From Table \ref{tbl:pipline}, our pipeline outperforms the SOTA when
evaluated in pipeline manner.
The results also emphasize the
importance of tokenization:  without gold tokenization
%\nascomment{changed ``segmentation'' to ``tokenization'' here, twice},
UD parsing
performance drops by about four points.
